---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
library(dplyr)
library(readr)
library(tidyr)
```

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

```{r}
# Transform Death columns to long form
deaths <- av %>%
  pivot_longer(cols = starts_with("Death"), 
               names_to = "Death_Time", 
               values_to = "Death") %>%
  mutate(Time = parse_number(Death_Time),
         Death = tolower(Death)) %>%
  select(-Death_Time)

# Transform Return columns to long form
returns <- av %>%
  pivot_longer(cols = starts_with("Return"), 
               names_to = "Return_Time", 
               values_to = "Return") %>%
  mutate(Time = parse_number(Return_Time),
         Return = tolower(Return)) %>%
  select(-Return_Time)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Calculate the average number of deaths per Avenger
# We understand that the code below only considers the avengers who have at least 1 death. We provide the second code that consider all avengers right below this code.
average_deaths <- deaths %>%
  filter(Death == "yes") %>%
  group_by(Name.Alias) %>%
  summarize(total_deaths = n()) %>%
  summarize(avg_deaths = mean(total_deaths, na.rm = TRUE))
```

```{r}
# Calculate the total number of deaths for each Avenger, including those with zero deaths
average_deaths <- deaths %>%
  mutate(Death = ifelse(Death == "yes", 1, 0)) %>%  # Convert "yes" to 1 and everything else to 0
  group_by(Name.Alias) %>%
  summarize(total_deaths = sum(Death, na.rm = TRUE)) %>%
  summarize(avg_deaths = mean(total_deaths, na.rm = TRUE))

# Output the result
average_deaths
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### Nhat Le

#### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.
My statement: “If a character is killed but then secretly hidden away in a stasis tube, they died.”

#### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
# Check for characters who have both a death and a return, implying a "hidden away" scenario
hidden_deaths <- deaths %>%
  filter(Death == "yes") %>%
  inner_join(returns %>% filter(Return == "yes"), by = c("Name.Alias", "Time")) %>%
  select(Name.Alias, Time, Death, Return)

# Display the first few rows to see examples of "hidden away" deaths
head(hidden_deaths)
```

```{r}
# Count the number of such scenarios
num_hidden_deaths <- hidden_deaths %>%
  nrow()
num_hidden_deaths
```

#### Include your answer

For this statement, we’ll interpret “killed” as a record of “yes” in the Death column in our dataset. However, since the concept of “hidden away in a stasis tube” implies they returned without a true resurrection, we’ll look for cases where a character has a recorded death followed by a recorded return in the Return column, regardless of how that return is defined.

According to the dataset, there are 67 instances where a character is recorded as having died and later returned. This supports the notion that even if a character was “hidden away” or returned through other means, their initial recorded “death” is considered legitimate in the dataset. This aligns with the statement from FiveThirtyEight that such characters are indeed counted as having “died,” even if they were later revived or returned from stasis.

### Pranava Sai Maganti
#### FiveThirtyEight Statement
The statement I am analyzing is: 

> If Iron Man fakes his death to chill out for a while but all his friends (and the reader) know he’s just kicking it somewhere, he didn’t die.

#### Include the code

```{r}
# Define Iron Man's aliases
alias1 <- "Anthony Edward \"Tony\" Stark"
alias2 <- "Anthony Edward Stark"

# Filter for each alias separately
iron_man_data_alias1 <- av %>% filter(Name.Alias == alias1)
iron_man_data_alias2 <- av %>% filter(Name.Alias == alias2)

# Select death and return columns for each alias
iron_man_deaths_alias1 <- iron_man_data_alias1 %>% select(starts_with("Death"))
iron_man_returns_alias1 <- iron_man_data_alias1 %>% select(starts_with("Return"))

iron_man_deaths_alias2 <- iron_man_data_alias2 %>% select(starts_with("Death"))
iron_man_returns_alias2 <- iron_man_data_alias2 %>% select(starts_with("Return"))

# Count the number of deaths and returns for each alias
iron_man_death_count_alias1 <- sum(iron_man_deaths_alias1 == "YES", na.rm = TRUE)
iron_man_return_count_alias1 <- sum(iron_man_returns_alias1 == "YES", na.rm = TRUE)
iron_man_no_return_count_alias1 <- sum(
  (iron_man_deaths_alias1$Death1 == "YES" & iron_man_returns_alias1$Return1 == "NO") |
  (iron_man_deaths_alias1$Death2 == "YES" & iron_man_returns_alias1$Return2 == "NO") |
  (iron_man_deaths_alias1$Death3 == "YES" & iron_man_returns_alias1$Return3 == "NO") |
  (iron_man_deaths_alias1$Death4 == "YES" & iron_man_returns_alias1$Return4 == "NO") |
  (iron_man_deaths_alias1$Death5 == "YES" & iron_man_returns_alias1$Return5 == "NO"),
  na.rm = TRUE
)

iron_man_death_count_alias2 <- sum(iron_man_deaths_alias2 == "YES", na.rm = TRUE)
iron_man_return_count_alias2 <- sum(iron_man_returns_alias2 == "YES", na.rm = TRUE)
iron_man_no_return_count_alias2 <- sum(
  (iron_man_deaths_alias2$Death1 == "YES" & iron_man_returns_alias2$Return1 == "NO") |
  (iron_man_deaths_alias2$Death2 == "YES" & iron_man_returns_alias2$Return2 == "NO") |
  (iron_man_deaths_alias2$Death3 == "YES" & iron_man_returns_alias2$Return3 == "NO") |
  (iron_man_deaths_alias2$Death4 == "YES" & iron_man_returns_alias2$Return4 == "NO") |
  (iron_man_deaths_alias2$Death5 == "YES" & iron_man_returns_alias2$Return5 == "NO"),
  na.rm = TRUE
)

# Display the counts and detailed death/return records for each alias
results <- list(
  "Alias 1 - Anthony Edward 'Tony' Stark" = list(
    "Death Count" = iron_man_death_count_alias1,
    "Return Count (Returned After Death)" = iron_man_return_count_alias1,
    "No Return Count (Died Without Returning)" = iron_man_no_return_count_alias1,
    "Death Records" = iron_man_deaths_alias1,
    "Return Records" = iron_man_returns_alias1
  ),
  "Alias 2 - Anthony Edward Stark" = list(
    "Death Count" = iron_man_death_count_alias2,
    "Return Count (Returned After Death)" = iron_man_return_count_alias2,
    "No Return Count (Died Without Returning)" = iron_man_no_return_count_alias2,
    "Death Records" = iron_man_deaths_alias2,
    "Return Records" = iron_man_returns_alias2
  )
)

results
```

#### Include your answer

##### Analysis and Results

Based on the results, the analysis of Iron Man’s death and return records across both aliases, "Anthony Edward 'Tony' Stark" and "Anthony Edward Stark," reveals distinct outcomes:

##### For "Anthony Edward 'Tony' Stark," we observe:
- **Death Count**: 1 recorded death.
- **Return Count**: 1 recorded return, indicating that Iron Man returned after his recorded death.
- **No Return Count**: 0, meaning there were no cases where he died without a subsequent return.

##### For "Anthony Edward Stark," we find:
- **Death Count**: 1 recorded death.
- **Return Count**: 0 recorded returns, signifying that he did not return after this death.
- **No Return Count**: 1, which confirms that there is a case where Iron Man died without a return.

##### Conclusion

These results support the statement that if Iron Man fakes his death but everyone knows he’s still alive or he later returns, it’s not considered a permanent death. For the alias "Anthony Edward 'Tony' Stark," Iron Man returned after his recorded death, aligning with this interpretation. However, for the alias "Anthony Edward Stark," he has one recorded death without a return, indicating that in this case, his death was not staged or temporary and could be seen as more definitive in the dataset's context.


### Manas Mathur

#### FiveThirtyEight Statement

> The quote I'm analyzing is: "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team."

#### Include the code

```{r}
av %>% 
  filter(Death1 == "YES") %>% 
  count()
```

#### Include your answer

I was able to conclude that the statement given that 69 avengers had died at least one time is true. This was done by checking the count of the value "Yes" in the 'Death1' column of av.

### Ananya Ramji

```{r}
deaths
```
```{r}
returns
```


#### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.
There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.


#### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
first_return <- deaths %>%
  filter(Death == "yes", Return1 == "YES", Time == 1)
first_did_not_return <- deaths %>%
  filter(Death == "yes", Return1 == "NO", Time == 1)

second_return <- deaths %>%
  filter(Death == "yes", Return2 == "YES", Time == 2)
second_did_not_return <- deaths %>%
  filter(Death == "yes", Return2 == "NO", Time == 2)

third_return <- deaths %>%
  filter(Death == "yes", Return3 == "YES", Time == 3)
third_did_not_return <- deaths %>%
  filter(Death == "yes", Return3 == "NO", Time == 3)

nrow(first_return)/(nrow(first_did_not_return)+nrow(first_return))
nrow(second_return)/(nrow(second_did_not_return)+nrow(second_return))
nrow(third_return)/(nrow(third_did_not_return)+nrow(third_return))
 
```

#### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

After completing the analysis and calculating the percentages of Avengers who returned after their first, second, and third death events, I found the following results:

The percentage of Avengers who returned after their first death was 66% or two-thirds.
The percentage of Avengers who returned after their second death was 50% or half.
The percentage of Avengers who returned after their third death was 50% or half.


## TEAM SUMMARY

Upload your changes to the repository. Discuss and refine answers as a team.
-- HELP ME SUMMARIZE THIS OR IF YOU GUYS FINISH. JUST LET ME KNOW AND I'LL JUST SUMMARIZE IT (NHAT LE)-- 
